Natural selection in the Health and Retirement Study

### 2023-09-16 hrs-selection.qmd
weight income residualization

calc_income <- function (x, eth) {
  x <- x |> filter(ethnicity == eth)
  ages <- x |> select(matches("r\\d+agey_e"))
  income <- x |> select(matches("r\\d+iearn"))
  income <- rowMeans(income, na.rm = TRUE)
  income_r <- resid(lm(income ~ factor(x$rabyear)))
  income_r <- income_r - min(income_r) + 1

@@

for (eth in eths) {
  pgen <- pgens[[eth]]
  names(pgen) <- shorten(names(pgen))
  pgs <- pgen |>
    select(-hhid, -pn, -version,
           -matches("^PC\\d"),
           -matches("^NEB"), # number ever born
           # outdated, or EDU3 without 23andMe:
           -EDU2, -EDU3, -HEIGHT, -BMI,
           -AD, -AD2, -GENCOG, -EVRSMK, -LONG, -CPD_TAG10,
           -ADHD_PGC10, -MDD
           ) |>
    names()

  rands[[eth]] <- inner_join(rand_orig, pgen,
                     join_by(hhid, pn),
                     unmatched = "drop",
                     relationship = "one-to-one")
}

>>>>

calc_income <- function (x, eth) {
  x <- x |> filter(ethnicity == eth)
  ages <- x |> select(matches("r\\d+agey_e"))
  income <- x |> select(matches("r\\d+iearn"))
  income <- rowMeans(income, na.rm = TRUE)
  income_r <- resid(lm(income ~ factor(x$rabyear), weights = x$weight))
  income_r <- income_r - min(income_r) + 1

@@

for (eth in eths) {
  pgen <- pgens[[eth]]
  names(pgen) <- shorten(names(pgen))
  rands[[eth]] <- inner_join(rand_orig, pgen,
                     join_by(hhid, pn),
                     unmatched = "drop",
                     relationship = "one-to-one")
}

pgs <- pgen4e |>
    rename_with(shorten) |> 
    # weird that I have to operate on whole df to use select semantics
    select(-hhid, -pn, -version,
           -matches("^PC\\d"),
           -matches("^NEB"), # number ever born
           # outdated, or EDU3 without 23andMe:
           -EDU2, -EDU3, -HEIGHT, -BMI,
           -AD, -AD2, -GENCOG, -EVRSMK, -LONG, -CPD_TAG10,
           -ADHD_PGC10, -MDD
           ) |>
    names()
